---
title: "Covid91 vaccine study Final2023"
author: "Leia Rodriguez"
output: 
  html_document:
    number_sections: true
    toc: true
    toc_depth: 5
date: "`r Sys.Date()`" 
---

```{r include=FALSE,echo=FALSE}
require(tidyverse)
require(tigerstats)
require(forcats)
FinalDat <-read.csv(file="final.csv",header=TRUE)
#recode to remove the annoying "X" from the dataframe
FinalData <- FinalDat  %>% select(-one_of("X"))
FinalData <- FinalData %>%
mutate(across(where(is.character), as.factor))
```

# Introduction

A new  preventative vaccine (covid921) is under test in clinical trials. Given the nature of this vaccine it is suspected that the drug may exhibit different effectiveness across the variety of subpopulations tracked. Consider population subgroups by gender( guys and gals), drug use, and sexual orientation. Write a complete report that separately considers this drug for each subgroup and your recommendations as to whether people in the subgroup should take the vaccine. Structure your report in four sections each devoted to one of the four subgroups, use null and alternative hypothesis statements and correct statistical verbage when referencing p-values in conclusions.

## Males

$H_0$ the covid-921 vaccine does not exhibit a significant difference in effectiveness between males 

$H_A$ the covid-921 vaccine does exhibit a significant difference in effectiveness between males  

```{r}
library(dplyr)
males = filter(FinalData, FinalData$sex=="M")
```

## Methods 

males is categorical and covid-921 is explanatory therefor we use cat~cat 

## Graphical results 

```{r}
barchartGC(~infected + treatment,data= males)
barchartGC(~infected + treatment,data= males, type="percent")

```

this bar graph shows that males who has covid and didnt take the vaccine did worse then the males that took the vaccine.The males that took the vaccine and did not have covid stayed failry the same to the ones that didnt get the vaccine.    

## Numerical results 

```{r}
table1 <- xtabs(~infected + treatment, data=males) 
rowPerc(table1)
colPerc(table1)
```
 
 this table shows that males who have covid that took the vaccine 30% still had covid and people who took the placebo, 69% still had covid who took the vaccine stayed fairly the same to those who didnt take the vaccine. 
 
## Inferental results 

## Females 

$H_0$ the covid-921 vaccine does not exhibit a significant difference in effectiveness between females 

$H_A$ the covid-921 vaccine does exhibit a significant difference in effectiveness between females  

```{r}
library(dplyr)
Female = filter(FinalData, FinalData$sex=="F")
```

## Methods 

female is categorical and covid-921 is explanatory therefor we use cat~cat

## Graphical results 

```{r}
barchartGC(~infected + treatment,data= Female)
barchartGC(~infected + treatment,data= Female, type="percent")
```

this bar graph shows that females who has covid and didnt take the vaccine did worse then the females that took the vaccine.The females that took the vaccine and did not have covid stayed failry the same to the ones that didnt get the vaccine.    

## Numerical results 

```{r}
table1 <- xtabs(~infected + treatment, data=Female) 
rowPerc(table1)
colPerc(table1)
```

this table shows that females who have covid that took the vaccine 40% still had covid and people who took the placebo, 60% still had covid who took the vaccine stayed fairly the same to those who didnt take the vaccine. 
 
## Inferental results 

## LGBTQ

$H_0$ the covid-921 vaccine does not exhibit a significant difference in effectiveness between LGBTQ

$H_A$ the covid-921 vaccine does exhibit a significant difference in effectiveness between LGBTQ  

```{r}
library(dplyr)
LGBTQ = filter(FinalData, FinalData$LGBTQ=="gay")
```

## Methods 

LGBTQ is categorical and covid-921 is explanatory therefor we use cat~cat

## Graphical results 

```{r}
barchartGC(~infected + treatment,data= LGBTQ)
barchartGC(~infected + treatment,data= LGBTQ, type="percent")
```

this bar graph shows that LGBTQ who has covid and didnt take the vaccine did better then the LGBTQ that took the vaccine.The LGBTQ that took the vaccine and did not have covid stayed failry the same to the ones that didnt get the vaccine. 

## Numerical results 

```{r}
table1 <- xtabs(~infected + treatment, data=LGBTQ) 
rowPerc(table1)
colPerc(table1)
```

this table shows that LGBTQ who have covid that took the vaccine 84% still had covid and people who took the placebo 15% still had covid. who took the vaccine stayed fairly the same to those who didnt take the vaccine.  

## Inferental results 


## Druggies 

$H_0$ the covid-921 vaccine does not exhibit a significant difference in effectiveness between drugies

$H_A$ the covid-921 vaccine does exhibit a significant difference in effectiveness between drugies  

```{r}
library(dplyr)
Druggies = filter(FinalData, FinalData$DrugUser=="D")
```

## Methods 

Druggies is categorical and covid-921 is explanatory therefor we use cat~cat

## Graphical results 

```{r}
barchartGC(~infected + treatment,data= Druggies)
barchartGC(~infected + treatment,data= Druggies, type="percent")
```

this bar graph shows the druggies had no reaction to taking the placebo or vacine and stayed failry the same. 

## Numerical results 

```{r}
table1 <- xtabs(~infected + treatment, data=Druggies) 
rowPerc(table1)
colPerc(table1)
```

This table shows that there was no results to the infected and not affect who took the placebo and vaccine. 

## Inferental results 

# Overall Results and Conclusions